---
title: 'Adapt into the CausalPath Format'
author: "Will Yashar"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: no
    theme: lumen
    toc: yes
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Packages
```{r}
library(tidyverse)
library(BBmisc)
```

# Load data
```{r}
data <- read.csv("data/combinedPhosphoData.tsv", sep='\t')
metadata <- read.csv("metadata.tsv", sep='\t')
```

# Make results directory
```{r}
dir.create(path = "results", showWarnings = FALSE)
```

# Create new columns
```{r}
# Remove duplicated rows
data <- data %>% distinct()

# Create a function to remove lower case letters (s, t, y) and replace with hyphens
letter_replace <- function(x) {
  if (x != "") {
      new <- str_split(x, "[a-z]")[[1]]
      new <- paste(new, collapse="-")
      new <- str_replace(new, pattern = "-$", replacement = "")
      return(new)
  } else {
    return(x)
  }
}

# Save as new column
site_letters_replaced <- lapply(data$site, letter_replace)
data$site_letters_replaced <- unlist(site_letters_replaced)

# Create a function that adds peptide or proteins to the site to make a unique ID
# This will become the ID column
unique_id <- function(x) {
  
  # Extract site and gene
  site <- data[x, "site_letters_replaced"]
  gene <- data[x, "gene"]
  
  # Identify what will be the reference for the ID
  # The reference will be the site, unless a site isn't present
  if (site != "") {
    ref <- site
  } else {
    ref <- gene}
  
  # Extract and format peptide
  peptide <- data[x, "peptide"]
  if(peptide == NaN) {
    peptide <- ""
  }
  
  # Extract and format protein
  protein <- data[x, "protein"]
  if(protein == NaN) {
    protein <- ""
  } else {
    protein <- str_split(protein, "_HUMAN")[[1]][1]
    protein <- paste(str_split(protein, "[|]")[[1]], collapse = "-")
  }
  
  # Concatenate and return unique ID
  new <- paste0(ref, peptide, protein, collapse="-")
  return(new)
}

# Save as new column
rows <- seq(1:nrow(data))
unique_id <- lapply(rows, unique_id)
data$unique_id <- unlist(unique_id)

# Create a function to remove gene names from sites
# Replace hyphens with pipe symbols
# This will become the Site column
# tests: 719, 875817, 2931649
remove_genes_add_pipes <- function(x) {
  
  # Extract site and gene
  site <- data[x, "site_letters_replaced"]
  gene <- data[x, "gene"]
  
  if (site != "") {
    # Remove gene
    no_genes <- str_split(site, paste0(gene, "-"))[[1]][2]
    
    # Replace hypthens with pipes
    no_genes_with_pipes <- str_replace_all(no_genes, pattern = "-", replacement = "|")
    
    return(no_genes_with_pipes)
  } else {
    
    return(site)
    
  }
  
}

# Save as new column
site_no_genes_with_pipes <- lapply(rows, remove_genes_add_pipes)
data$site_no_genes_with_pipes <- unlist(site_no_genes_with_pipes)
```

# Select and rearrange
```{r}
# Add an Effect column
data$Effect <- ""

# Rename and rearrange columns
prep_cp_format <- data %>% 
  rename(ID = unique_id,
         Sites = site_no_genes_with_pipes,
         Symbols = gene) %>% 
  select(ID, Sites, Effect, Symbols, sample, logRatio)

# Replace NaN and NA values
prep_cp_format$ID[( (is.nan(prep_cp_format$ID)) | (is.na(prep_cp_format$ID)) )] <- ""
prep_cp_format$Sites[( (is.nan(prep_cp_format$Sites)) | (is.na(prep_cp_format$Sites)) )] <- ""
prep_cp_format$Symbols[( (is.nan(prep_cp_format$Symbols)) | (is.na(prep_cp_format$Symbols)) )] <- ""

# Remove duplicated rows
prep_cp_format <- prep_cp_format %>% distinct()

# Pivot wider
cp_format <- pivot_wider(prep_cp_format, 
                         id_cols = c(ID, Sites, Effect, Symbols),
                         names_from = sample, 
                         values_from = logRatio)

# Replace empty values
cp_format[is.na(cp_format)] <- NaN

# Save as a data.frame
cp_format <- as.data.frame(cp_format)
write.table(cp_format, file = "aml_data_cp_format_unnormalized.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
```

# Pre-Normalization PCA
# Workflow: http://juliejosse.com/wp-content/uploads/2018/05/DataAnalysisMissingR.html#regression_with_na_(quantitative)
```{r}
library(naniar)
library(VIM)
library(missMDA)
library(ggbiplot)
library(ggthemes)

sparse_pca <- function(data, data_type) {
  
  # Extract sample values
  sample_vals <- data[,5:ncol(data)]
  
  # Visualize missing values (# of missing values)
  png(sprintf("results/%s_n_missing_values.png", data_type))
  gg_miss_var(sample_vals)
  dev.off()
  
  # Visualize missing values (% of missing values)
  png(sprintf("results/%s_n_missing_values.png", data_type))
  res <- summary(aggr(sample_vals, sortVar=TRUE))$combinations
  dev.off()
  
  # Identify most sparse combinations
  head(res[rev(order(res[,2])),])
  
  # Impute dataset in preparation for PCA
  imputed <- imputePCA(sample_vals, ncp = 2)
  imputed$completeObs[1:3,]
  
  # Impute using a random forest implementatino
  library(missForest)
  imputed <- missForest(sample_vals)
  imputed$ximp
  imputed$OOBerror
  
  # Perform PCA
  pca <- PCA(imputed, quanti.sup = 1, quali.sup = 12, ncp = 2, graph=FALSE)
  
  # Visualize pca
  plot(pca, hab=12)
  
  # Visualize PCA
  ggbiplot(pca) + theme_clean()
  
  # Visualize PCA
  ggbiplot(pca, ellipse = TRUE, var.axes=FALSE) + theme_clean()
  
}

# Extract sample values
cp_vals <- cp_format[,5:ncol(cp_format)]

# Visualize missing values (# of missing values)
png("results/pre-normalization_n_missing_values.png")
gg_miss_var(cp_vals)
dev.off()

# Visualize missing values (% of missing values)
png("results/pre-normalization_n_missing_values.png")
res <- summary(aggr(cp_vals, sortVar=TRUE))$combinations
dev.off()

# Identify most sparse combinations
head(res[rev(order(res[,2])),])

# Impute dataset in preparation for PCA
cp_vals_imputed <- imputePCA(cp_vals, ncp = 2)
cp_vals_imputed$completeObs[1:3,]

# Perform PCA
pca <- PCA(imp, quanti.sup = 1, quali.sup = 12, ncp = nb$ncp, graph=FALSE)

# Visualize pca
plot(res.pca, hab=12)

# Visualize PCA
ggbiplot(pre_norm_pca) + theme_clean()

# Visualize PCA
ggbiplot(pre_norm_pca, ellipse = TRUE, var.axes=FALSE) + theme_clean()
```


# Normalize

```{r}

# Pre-Normalization PCA
sparse_pca(cp_format, "pre-normalization")

library(quantable)

# Column normalize
cp_format_col_norm <- apply(cp_format[,5:ncol(cp_format)], MARGIN = 1, FUN = function(x) robustscale(x, dim = 2, center = TRUE, scale = TRUE, preserveScale = TRUE))

# Post-normalization PCA
sparse_pca(cp_format_col_norm, "col-normalized")
write.table(cp_format, file = "aml_data_cp_format_col_norm.tsv", sep = "\t", row.names = FALSE, quote = FALSE)

# Row normalize 
cp_format_row_norm <- apply(cp_format[,5:ncol(cp_format)], MARGIN = 2, FUN = function(x) robustscale(x, dim = 2, center = TRUE, scale = TRUE, preserveScale = TRUE))

# Post-normalization PCA
sparse_pca(cp_format_col_norm, "col-normalized")
write.table(cp_format, file = "aml_data_cp_format_col_norm.tsv", sep = "\t", row.names = FALSE, quote = FALSE)
```

# Harmonize metadata
```{r}
# Control
metadata$treatment[metadata$treatment == "No treatment"] <- "none"
metadata$treatment[metadata$treatment == "no treatment"] <- "none"
metadata$treatment[metadata$treatment == "DMSO"] <- "none"

# Trametinib
metadata$treatment[metadata$treatment == "trametinib"] <- "Trametinib"

# Ruxolitnib
metadata$treatment[metadata$treatment == "ruxolitnib"] <- "Ruxolitnib"

# Dasatinib
metadata$treatment[metadata$treatment == "dasatinib"] <- "Dasatinib"

# Sorafenib
metadata$treatment[metadata$treatment == "sorafenib"] <- "Sorafenib"

# Trametinib Withdrawn
metadata$treatment[metadata$treatment == "Trametinib Withdrawn"] <- "Trametinib_Withdrawn"
```

# Identify comparison groups
# MOLM-13
```{r}
# Store experiments
exptsList <- list()

## Cell Line
cellLine <- "MOLM-13"

### Control
tx <- "none"
values <- metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "sample"]
exptsList[[cellLine]][[tx]] <- list(values)

### Trametinib
tx <- "Trametinib"
timePoints <- unique(metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}

### MCP-1
tx <- "MCP-1"
timePoints <- unique(metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}

### Trametinib+MCP-1
tx <- "Trametinib+MCP-1"
timePoints <- unique(metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}
```

# Identify comparison groups
# MOLM-14-a
```{r}
## Cell Line
cellLine <- "MOLM14-a"

### Control
tx <- "none"
values <- metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "sample"]
exptsList[[cellLine]][[tx]] <- list(values)

### Gilteritinib
tx <- "Gilteritinib"
timePoints <- unique(metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}
```

# Identify comparison groups
# MOLM-14-b
```{r}
## Cell Line
cellLine <- "MOLM14-b"

### Control
tx <- "none"
values <- metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "sample"]
exptsList[[cellLine]][[tx]] <- list(values)

### Gilteritinib
tx <- "Gilteritinib"
timePoints <- unique(metadata[( (metadata$cellLine == cellLine) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}
```

# Identify comparison groups
# MOLM-14
```{r}
## Cell Line
cellLine <- "MOLM14"

### Control
tx <- "none"
values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) ), "sample"]
exptsList[[cellLine]][[tx]] <- list(values)

### Gilteritinib
tx <- "Gilteritinib"
timePoints <- unique(metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) ), "timePoint"])

for (time in timePoints) {
  values <- metadata[( (grepl(cellLine, metadata$cellLine) == TRUE) & (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
  exptsList[[cellLine]][[tx]][time] <- list(values)
}
```

# Create parameters file
```{r}

for (cell in names(exptsList)) {

  for (tx in names(exptsList[[cell]])) {

    for (time in names(exptsList[[cell]][[tx]])) {

      # create directory
      dir_path <- sprintf("results/%s/%s/%s", cell, tx, time)
      dir.create(path = dir_path, showWarnings = FALSE, recursive = TRUE)

      # Set parameters
      parameters = c("proteomics-values-file = aml_data_cp_format.tsv",
                 "id-column = ID",
                 "symbols-column = Symbols",
                 "sites-column = Sites",
                 "effect-column = Effect",
                 "value-transformation = significant-change-of-mean",
                 "fdr-threshold-for-data-significance = 0.05 phosphoprotein",
                 "fdr-threshold-for-data-significance = 0.05 protein",
                 "calculate-network-significance = true",
                 "permutations-for-significance = 100",
                 "generate-data-centric-graph = false",
                 "show-insignificant-data = false",
                 "color-saturation-value = 50",
                 "custom-resource-directory = /Users/yashar/gitPackages/causalpath/.panda",
                 "use-network-significance-for-causal-reasoning = false")

      # Set control values
      i <- length(parameters)
      for (control in unlist(exptsList[[cell]][['none']][[1]])) {
        i <- i + 1
        parameters[i] <- paste0("control-value-column = ", control)
      }

      # Set test columns
      for (test in unlist(exptsList[[cell]][[tx]][time])) {
        i <- i + 1
        parameters[i] <- paste0("test-value-column = ", test)
      }

      # Write to parameters file
      file_path <- file(sprintf("%s/parameters.txt", dir_path))
      writeLines(parameters, file_path)
      close(file_path)

      # Save Causal Path File
      write.table(cp_format, file = sprintf("%s/aml_data_cp_format.tsv", dir_path), sep = "\t", row.names = FALSE, quote = FALSE)

    }
    
  }
  
}
```

# All Iterations
# Pool all controls together
```{r}
# Store experiments
exptsList <- list()

## Cell Line
cellLine <- "all"

### Control
tx <- "none"
values <- metadata[(metadata$treatment == tx), "sample"]
exptsList[[cellLine]][[tx]] <- list(values)

### All treatments
txs <- unique(metadata[, "treatment"])
txs <- txs[( (txs != "none") & (txs != NaN) )]

for (tx in txs) {
  
  timePoints <- unique(metadata[( (metadata$treatment == tx) ), "timePoint"])

  for (time in timePoints) {
    values <- metadata[( (metadata$treatment == tx) & (metadata$timePoint == time) ), "sample"]
    exptsList[[cellLine]][[tx]][time] <- list(values)
  }
  
}

```

# Create parameters file
```{r}

for (cell in names(exptsList)) {

  for (tx in names(exptsList[[cell]])) {

    for (time in names(exptsList[[cell]][[tx]])) {

      # create directory
      dir_path <- sprintf("results/%s/%s/%s", cell, tx, time)
      dir.create(path = dir_path, showWarnings = FALSE, recursive = TRUE)

      # Set parameters
      parameters = c("proteomics-values-file = aml_data_cp_format.tsv",
                 "id-column = ID",
                 "symbols-column = Symbols",
                 "sites-column = Sites",
                 "effect-column = Effect",
                 "value-transformation = significant-change-of-mean",
                 "fdr-threshold-for-data-significance = 0.05 phosphoprotein",
                 "fdr-threshold-for-data-significance = 0.05 protein",
                 "calculate-network-significance = true",
                 "permutations-for-significance = 100",
                 "generate-data-centric-graph = false",
                 "show-insignificant-data = false",
                 "color-saturation-value = 50",
                 "custom-resource-directory = /Users/yashar/gitPackages/causalpath/.panda",
                 "use-network-significance-for-causal-reasoning = false")

      # Set control values
      i <- length(parameters)
      for (control in unlist(exptsList[[cell]][['none']][[1]])) {
        i <- i + 1
        parameters[i] <- paste0("control-value-column = ", control)
      }

      # Set test columns
      for (test in unlist(exptsList[[cell]][[tx]][time])) {
        i <- i + 1
        parameters[i] <- paste0("test-value-column = ", test)
      }

      # Write to parameters file
      file_path <- file(sprintf("%s/parameters.txt", dir_path))
      writeLines(parameters, file_path)
      close(file_path)

      # Save Causal Path File
      write.table(cp_format, file = sprintf("%s/aml_data_cp_format.tsv", dir_path), sep = "\t", row.names = FALSE, quote = FALSE)

    }
    
  }
  
}
```